CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to Support Active Travel}

### 2023-10-17 cover-letter.qmd
Updates for ETRR

We are submitting our manuscript entitled "CRUSE to Safe Cycling in Ireland: An Open Source Tool to Support Active Travel" for consideration for publication in the European Transport Research Review.
It has a word count of 4,721, excluding references.

This manuscript describes the development of an open source tool to support active travel in Ireland.
The paper describes the design, features and potential uses of the CRUSE Tool.
We believe that the paper is important because it marks a substantial step forward in the development of open source tools to support active travel.
The work presented goes substantially beyond previously published research such as the Propensity to Cycle Tool for England and Wales.

@@

Robin Lovelace on behalf of all authors, Joey Talbot, Eugeni Vidal-Tortosa, and Hussein Mahfouz (University of Leeds), Elaine Brick, Peter Wright, Gary O'Tool (AECOM), Dan Brennan and Suzanne Meade (Transport Infrastructure Ireland)

>>>>

We are submitting our manuscript entitled "CRUSE to Safe Cycling in Ireland: An Open Source Tool to Support Active Travel" for consideration for publication in the European Transport Research Review, as part of a **Full Paper Submission for TRA 2024 (note: this option was not available the dropdown menu in the Thematic Series question, please ensure it is in the correct category)**.
It has a word count of 4,721, excluding references.

Please see the full list of authors and the abstract in the title page below.
The manuscript describes the development of an open source tool to support active travel in Ireland.
The paper describes the design, features and potential uses of the CRUSE Tool.
We believe that the paper is important because it marks a substantial step forward in the development of open source tools to support active travel.
The work presented goes substantially beyond previously published research such as the Propensity to Cycle Tool for England and Wales.

@@

Robin Lovelace on behalf of all authors:

Robin Lovelace (University of Leeds, r.lovelace@leeds.ac.uk)

Joey Talbot (University of Leeds, J.D.Talbot@leeds.ac.uk)

Eugeni Vidal-Tortosa (University of Leeds, E.VidalTortosa@leeds.ac.uk)

Hussein Mahfouz (University of Leeds, tshma@leeds.ac.uk)

Suzanne Meade (Transport Infrastructure Ireland, Suzanne.Meade@tii.ie)

Dan Brennan (Transport Infrastructure Ireland, Dan.Brennan@tii.ie)

Peter Wright (AECOM, peter.wright1@aecom.com)

Gary O'Toole (AECOM, Gary.OToole@aecom.com)

Elaine Brick (AECOM, elaine.brick@aecom.com)

# Declarations

## Availability of data and material {.unnumbered}

Data was obtained from the Central Statistics Office (CSO) and Transport Infrastructure Ireland (TII) under license and cannot be shared publicly.
The code used to generate the results presented in this paper is available <!-- at https://github.com/cruse-bike. --> \[To be added after paper is accepted.\]

## Funding {.unnumbered}

The work was funded by Transport Infrastructure Ireland (TII).

## Acknowledgements {.unnumbered}

Thanks to Paul MacDonald and Donal Hodgins (Kildare County Council), for testing early versions of the tool and for their input into the project.
Thanks to Ciaran Maguire, Catherine Swift, and others at AECOM for their input into the project.
Thanks to Martin Lucas-Smith and Simon Nutall from CycleStreets for their input into the project.

## Competing interests {.unnumbered}

The authors declare that they have no competing interests.

## Authors' contributions {.unnumbered}

RL led the development of the CRUSE Tool, with input from JT, EV, HM, EB, PW, GOT, DB, and SM.
SM instigated the project and coordinated its policy impact.
PW, EB and GOT provided project management and transport engineering expertise.

### 2023-10-17 paper.qmd
Updates for ETRR

# # Uncomment for draft versions:
format: 
  html: default
  # docx: default
# # Uncomment for formatted version
#   elsevier-pdf:
#     linestretch: 2
#     keep-tex: true
#     journal:
#       name: European Transport Research Review
#       formatting: preprint
#       model: 3p
#       cite-style: super
# # #   # pdf: default

# # # Uncomment the following line to produce the cover page
# author:
#   - name: Robin Lovelace
#     email: r.lovelace@leeds.ac.uk
#     corresponding: true
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Joey Talbot
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Eugeni Vidal-Tortosa
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Hussein Mahfouz
#     affiliations:
#       - name: University of Leeds
#         address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
#   - name: Elaine Brick
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Peter Wright
#     affiliations:
#       - name: AECOM
#         address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
#   - name: "Gary O'Tool"
#     affiliations:
#       - name: AECOM
#         address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
#   - name: Dan Brennan
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
#   - name: Suzanne Meade
#     affiliations:
#       - name: Transport Infrastructure Ireland
#         address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  Legislation mandates improved road safety and decarbonisation of predominantly road-based transport network in Ireland and other countries. Increasing cycling transport, making it safer and more attractive, is essential to this end. Cycling in Ireland represents only 3% of total modal share as of the latest Census (2016), but accounts for 20% of serious injuries and 7% of all fatalities. A planning tool based on high resolution datasets on baseline and potential cycling levels could greatly assist in achieving this goal.

@@

Data was obtained from the Central Statistics Office (CSO) and Transport Infrastructure Ireland (TII) under license and cannot be shared publicly.
The code used to generate the results presented in this paper is available <!-- at https://github.com/cruse-bike. --> \[To be added after paper is accepted.\]

@@

The work was funded by Transport Infrastructure Ireland (TII).

@@

Thanks to Paul MacDonald and Donal Hodgins (Kildare County Council), for testing early versions of the tool and for their input into the project.
Thanks to Ciaran Maguire, Catherine Swift, and others at AECOM for their input into the project.
Thanks to Martin Lucas-Smith and Simon Nutall from CycleStreets for their input into the project.

## Competing interests {.unnumbered}

The authors declare that they have no competing interests.

## Authors' contributions {.unnumbered}

\[To be added after paper is accepted.\]

```{=html}
<!-- RL led the development of the CRUSE Tool, with input from JT, EV, HM, EB, PW, GOT, DB, and SM.
SM instigated the project and coordinated its policy impact.
PW, EB and GOT provided project management and transport engineering expertise. -->
```

>>>>

# # Uncomment for draft versions:
format: 
  # html: default
  # docx: default
# # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # #   # pdf: default

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY, UK
  - name: "Gary O'Tool"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  Legislation mandates improved road safety and decarbonisation of predominantly road-based transport network in Ireland and other countries. Increasing cycling transport, making it safer and more attractive, is essential to this end. Cycling in Ireland represents only 3% of total modal share as of the latest Census (2016), but accounts for 20% of serious injuries and 7% of all fatalities. A planning tool based on high resolution datasets on baseline and potential cycling levels could greatly assist in achieving this goal.

@@

Data was obtained from the Central Statistics Office (CSO) and Transport Infrastructure Ireland (TII) under license and cannot be shared publicly.
The code used to generate the results presented in this paper is fully reproducible and available at \[to be added post review.\]

@@

This work was funded by Transport Infrastructure Ireland (TII).

@@

\[To be added post review.\]

### 2023-10-17 paper.tex
Updates for ETRR

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to
Support Active Travel} }

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Data was obtained from the Central Statistics Office (CSO) and Transport
Infrastructure Ireland (TII) under license and cannot be shared
publicly. The code used to generate the results presented in this paper
is available {[}To be added after paper is accepted.{]}

@@

Thanks to Paul MacDonald and Donal Hodgins (Kildare County Council), for
testing early versions of the tool and for their input into the project.
Thanks to Ciaran Maguire, Catherine Swift, and others at AECOM for their
input into the project. Thanks to Martin Lucas-Smith and Simon Nutall
from CycleStreets for their input into the project.

\hypertarget{competing-interests}{%
\subsection*{Competing interests}\label{competing-interests}}
\addcontentsline{toc}{subsection}{Competing interests}

The authors declare that they have no competing interests.

\hypertarget{authors-contributions}{%
\subsection*{Authors' contributions}\label{authors-contributions}}
\addcontentsline{toc}{subsection}{Authors' contributions}

{[}To be added after paper is accepted.{]}

>>>>

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to
Support Active Travel} }
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

\author[1]{Eugeni Vidal-Tortosa%
%
}

\author[1]{Hussein Mahfouz%
%
}

\author[2]{Elaine Brick%
%
}

\author[3]{Peter Wright%
%
}

\author[2]{Gary O'Tool%
%
}

\author[4]{Dan Brennan%
%
}

\author[4]{Suzanne Meade%
%
}

\affiliation[1]{organization={University of
Leeds},addressline={Institute for Transport Studies, Leeds, LS2 9JT,
United Kingdom},postcodesep={}}
\affiliation[2]{organization={AECOM},addressline={Unit 6, Galway
Technology Park, Parkmore, Galway, Ireland},postcodesep={}}
\affiliation[3]{organization={AECOM},addressline={Winslade House,
Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY,
UK},postcodesep={}}
\affiliation[4]{organization={Transport Infrastructure
Ireland},addressline={Parkgate Business Centre, Parkgate Street, Dublin
8, D08 DK10, Ireland},postcodesep={}}

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

Data was obtained from the Central Statistics Office (CSO) and Transport
Infrastructure Ireland (TII) under license and cannot be shared
publicly. The code used to generate the results presented in this paper
is fully reproducible and available at {[}to be added post review.{]}

@@

{[}To be added post review.{]}
